Anthropogenic plant communities of the Iberian Atlantic ecoregion

### 2023-12-21 doc/manuscript.Rmd
results section

To identify anthropogenic relevés in the SIVIM database we followed two complementary approaches. Fist, we used the expert system created to classify vegetation plots into EUNIS pan-European habitat classification types [@RN5061]. The expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species. We used the habitats assigned by the expert system to retrieve plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation). Second, we also retrieved relevés if they had been classified by the original authors of the relevé into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**). This produced an initial pool of 3,160 relevés from SIVIM, to which we added 89 relevés of urban or peri-urban plant communities conducted by us or extracted by us from the local literature [@RN5377; @RN5378]. After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652 outlier relevés which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion. We removed these, leaving a dataset of 2,508 relevés for further analysis: 2,419 relevés from SIVIM, originally recorded in 89 publications see (**Supplementary 2**); plus 89 relevés added by us.

## Definition of target alliances and training dataset

As a previous step before classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371]. The provisional list included 38 alliances.

Then, we filtered our dataset to include only those relevés which had been assigned by the authors to any of these 38 alliances (n = 2,201 relevés). We used sequential TWINSPANs to do a preliminary re-classification of the 2,201 filtered relevés, attempting to re-match the relevés into the *a priori* defined checklist of 38 alliances. We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' [@RN5380] to do the TWINSPANs. To start, we did a TWINSPAN classification into 15 clusters, followed by separate TWINSPANS within each cluster. This iterative process allowed us to validate 28 alliances and assign 1,725 relevés to them, leaving 783 relevés as unclassified.

## Semi-supervised classification

Next, we applied a semi-supervised classification to assign the so far unclassified relevés to one of the validated 28 alliances. Semi-supervised classification uses previously defined groups of data to assign the unclassified relevés to them (as in supervised classification) but also creates new groups to place the data points that do not match the already existing groups (as in unsupervised classification), allowing to detect new groups and to update the previous classification [@RN5383]. We used the 1,719 relevés that we had already re-classified into the 28 alliances as a training set for the semi-supervised classification. Attempts by the classification to create new groups resulted on new groups with no ecological significance, and thus we kept the results of the classification into the 28 *a priori* alliances. For the classification, we applied a noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional relevés [@RN5382]. We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional relevés; and we set the distance to the noise class to d = 1. We performed this analysis with the R package 'vegclust' [@RN5383]. The semi-supervised classification with noise clustering resulted in the final classification of 2,081 relevés into 28 alliances and 427 relevés left in the noise group. We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to detect the maximum number of mutually different clusters in the filtered dataset; and Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the vegetation groups.

## Characteristic species

We calculated sets of characteristic species [@RN5061] for each of the 28 alliances. We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the alliance; constant species as species with a frequency higher than 50% in the alliance; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.

## Community traits and ecological requirements

To describe the resulting 28 alliances, we compiled a dataset of species traits and ecological preferences for the species pool of the 2,508 classified relevés. For each species, we extracted from *Flora iberica* [-@RN3243] its maximum height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not. From FloraVeg.EU [-@RN5385] we extracted the species’ life forms and kept 'therophyte' and 'geophyte' for further analysis since these two categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package 'FactoMineR' [@RN3166]. We also extracted the ecological indicator values [@RN5384] and the disturbance indicator values [@RN5101] of each species. In the case of the ecological and disturbance indicators, we assigned values to species lacking them by (1) calculating weighted average values of the species with values for each plot (weighting by species cover), and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover). Then, for each relevé we calculated its community-weighted mean (weighting by species cover) for each trait and ecological preference, and explored the main axes of variation using PCA, separately for species traits and ecological preferences.

>>>>

To identify anthropogenic relevés in the SIVIM database we followed two complementary approaches. Fist, we used the expert system created to classify vegetation plots into EUNIS pan-European habitat classification types [@RN5061]. The expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species. We used the habitats assigned by the expert system to retrieve plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation). Second, we also retrieved relevés if they had been classified by the original authors of the relevé into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**). This produced an initial pool of 3,160 relevés from SIVIM, to which we added 89 relevés of urban or peri-urban plant communities conducted by us or extracted by us from the local literature [@RN5377; @RN5378]. After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652 outlier relevés which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion. We removed these, leaving a dataset of 2,508 relevés for the analyses of this article: 2,419 relevés from SIVIM, originally recorded in 89 publications see (**Supplementary 2**); plus 89 relevés added by us.

## Statistical analyses

[Definition of target alliances and training dataset.]{.underline} As a previous step before classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371]. Then, we filtered our dataset to include only those relevés which had been assigned by the authors to any of these alliances, and used sequential TWINSPANs to do a preliminary re-classification of these filtered relevés, attempting to re-match the relevés into the *a priori* defined checklist of alliances. We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' [@RN5380] to do the TWINSPANs. To start, we did a TWINSPAN classification into 15 clusters, followed by separate TWINSPANS within each cluster. This preliminary TWINSPAN allowed us to (1) validate which alliances were present in the study ecoregion; and (2) prepare a training dataset for semi-supervised classification (see below).

[Semi-supervised classification.]{.underline} Next, we applied a semi-supervised classification to the whole dataset. Semi-supervised classification uses a training subset of *a priori* classified relevés to classify a secondary subset of unclassified relevés [@RN5383]. Our training subset included all the relevés that (1) had been classified by the authors of the relevé into any of the alliances that we had validated during the previous TWINSPANs, and (2) had also been classified by us into these alliances during our TWINSPANs. The secondary subset included all the relevés which did not meet these two criteria. Our goal was to refine and not to simply extend the classification, so we applied the semi-supervised classification to the whole dataset, i.e. we allowed relevés from the training subset to be re-assigned to other groups during the classification. In addition, semi-supervised classification can create new groups to place data points that do not match the already *a priori* groups. Furthermore, we applied a noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional relevés [@RN5382]. We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional relevés; and we set the distance to the noise class to d = 1. We performed this analysis with the R package 'vegclust' [@RN5383]. We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to test the significant of the final vegetation groups; and Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the groups.

[Characteristic species and EUNIS habitat regionalization.]{.underline} We calculated sets of characteristic species [@RN5061] for each of the final vegetation groups. We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations. Additionally, we assigned to each group a regionalized level 4 EUNIS habitat code, using already codes when they existed, and creating new codes when not.

[Community traits and ecological requirements.]{.underline} To describe the resulting vegetation groups, we compiled a dataset of species traits and ecological preferences for the species pool of the classified relevés. For each species, we extracted from *Flora iberica* [-@RN3243] its maximum height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not. From FloraVeg.EU [-@RN5385] we extracted the species’ life forms and kept 'therophyte' and 'geophyte' for further analysis since these two categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package 'FactoMineR' [@RN3166]. We also extracted the ecological indicator values [@RN5384] and the disturbance indicator values [@RN5101] of each species. In the case of the ecological and disturbance indicators, we assigned values to species lacking them by (1) calculating weighted average values of the species with values for each plot (weighting by species cover), and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover). Then, for each relevé we calculated its community-weighted mean (weighting by species cover) for each trait and ecological preference, and explored the main axes of variation using PCA, separately for species traits and ecological preferences.

# Results

## Overview of the classification

Our revision of the literature produced a provisional list of 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion (**Supplementary 3**). In our study dataset (n = 2,508 relevés), there were 2,201 relevés that had been assigned by the authors to any of these alliances. We used these 2,201 relevés for a preliminary re-classification based on TWINSPAN, attempting to re-match the relevés into the *a priori* defined checklist of 38 alliances. This procedure allowed us to validate 28 alliances as present in the ecoregion. A total of 1,725 relevés had been classified by both the original authors and our TWINSPAN into these alliances, and these became our training subset for semi-supervised classification. Attempts by the semi-supervised classification to create new groups resulted on new groups with no ecological significance, and thus we kept the the 28 alliances as final vegetation groups (**Table 1**). Most of the alliances had a wide distribution in the ecoregion according to our dataset (**Figure 1**), although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories. The regional literature included in our dataset recognized 72 associations (**Supplementary 4**) within these alliances. The semi-supervised classification with noise clustering resulted in the final classification of 2,081 relevés into 28 alliances and 427 relevés left in the noise group.

Focusing on the classified plots, the dataset included 1,167 taxa or taxa aggregates. Of these, 101 were neophytes (i.e. 9% of the species pool). Regarding life forms, the species pool was dominated by hemicryptophytes (43%) and therophytes (41%). The 10 most frequent species were *Ochlopoa annua* (621 occurences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332) and *Anisantha sterilis* (301). Considering only those relevés with the most frequent plot size (20 m^2^, n = 256 relevés), the average species richness per plot was 16 (minimum = 3, maximum = 37).

DCA ordination of the floristic composition of the classified relevés (**Fig. 2**) suggested a high consistency of the vegetation classes *Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea* and *Bidentetea*. The alliances belonging to the two perennial ruderal vegetation classes (*Artemisietea vulgaris* and *Epilobietea angustifolii*) were mixed together. The winter-annual class *Chenopodietea* also showed high heterogeneity, specially because of the isolation of the altered dune alliance *Linario polygalifoliae-Vulpion alopecuri*.

## Community-level traits and ecological preferences

PCA ordination of the community-weighted means of the species traits (**Fig. 3A**) indicated that the first two axes explained 65% of the variability. The major contributors to axis 1 (41% variance explained) were community height, proportion of geophytes, mean month of flowering and proportion of therophytes; it separated taller communities with more geophytes and later flowering (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*) from shorter communities with more therophytes and earlier flowering (classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis* and *Sisymbrietea*). The major contributors to axis 2 (24% variance explained) were proportion of neophytes and width of the flowering season; it separated communities with more neophytes and longer flowering seasons (classes *Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris* and *Polygono-Poetea annuae*) from communities with less neophytes and shorter flowering seasons (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Chenopodietea*).

PCA ordination of the community-weighted means of the species ecological and disturbance preferences (**Fig. 3B**) indicated that the first two axes explained 56% of the variability. The major contributors to axis 1 (35% variance explained) were diturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities prefering less frequent disturbances and more shade (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*). The major contributors to axis 2 (21% variance explained) were nutrient requirements and disturbance severity; it separated more nitrophilous communities adapted to more severe disturbances (classes *Digitario-Eragrostietea*, *Sisymbrietea* and *Epilobietea angustifolii*) from less nitrophilous communities adapted to lower intensity disturbances (classes *Cymbalario-Parietarietea diffusae*, *Artemisietea vulgaris* and *Chenopodietea*).

## Classified anthropogenic plant communities

***Cymbalario-Parietarietea diffusae*** Chasmophytic vegetation of human-made walls.

1. *Galio valantiae-Parietarion judaicae* Vegetation of dry walls. Occupies warmer, drier and more nitriphicated situations than the other alliance in the class, has a shorter flowering season and is more adept to disturbances (**Fig. 4**). Wide distrubution in the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Parietaria judaica* aggr., *Centranthus ruber*, *Asplenium trichomanes*.

2. *Cymbalario-Asplenion* Vegetation of mesic walls. Occupies colder, wetter and less nitriphicated situations than the other alliance in the class, has a longer flowering season and is less adept to disturbances (**Fig. 4**). Wide distrubution in the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Cymbalaria muralis*.

***Polygono-Poetea annuae*** Dwarf vegetation of heavily-trampled sites.

3. *Polycarpion tetraphylli* Dwarf-annual trampled vegetation of warm and sunny sites. Occupies warmer, sunnier and less nitriphicated situations than the other alliances in the class and has an earlier flowering season (**Fig. 4**). Wide distrubution in the ecoregion, specially along the coast and in the more Mediterranean valleys of the interior (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Sagina apetala*, *Plantago coronopus*, *Crassula tillaea*, *Spergularia marina*.

4. *Polygono-Coronopodion* Dwarf-annual trampled vegetation of dry sites. Occupies colder, drier and more nitriphicated situations than the other alliances in the class, and has a higher proportion of neophytes (**Fig. 4**). Wide distrubution in the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Matricaria suaveolens*, *Polygonum aviculare*.

5. *Saginion procumbentis* Dwarf trampled vegetation of mesic sites. Occupies colder, wetter and more nitriphicated situations than the other alliance in the class, and is less dominated by therophytes (**Fig. 4**). Distribution concentrated in the central part of the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Sagina procumbens*, *Bryum argenteum*.

***Papaveretea rhoeadis*** Annual segetal vegetation of crops and gardens.

6. *Caucalidion lappulae* Annual segetal vegetation of cereal crops on base-rich soils. Adapted to more frequent and severe disturbances than the other alliances in the class  (**Fig. 4**). Distribution mostly limited to the more Mediterranean valleys south of the Cantabrian Mountains (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Cyanus segetum*, *Papaver rhoeas*, *Trifolium arvense*, *Anacyclus clavatus*, *Valerianella eriocarpa*.

7. *Scleranthion annui* Annual segetal vegetation of cereal crops on base-poor soils. Occupies more acidic soils than the other alliances in the class, and flowers earlier (**Fig. 4**). Wide distribution in the ecoregion, but specially in the west, where acidic bedrocks dominate (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Mibora minima*, *Rumex acetosella*.

8. *Oxalidion europeae* Annual segetal vegetation of gardens and root crops. Occupies wetter, shadier and more nitriphicated situations than the other alliances in the class,  has a longer flowering season and a higher proportion of neophytes (**Fig. 4**). Wide distribution in the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Stellaria media*, *Veronica persica*.

***Digitario sanguinalis-Eragrostietea minoris*** Summer-annual C4 segetal vegetation.

9. *Spergulo arvensis-Erodion cicutariae* Summer-annual C4 segetal vegetation. Late-flowering alliance rich in neophytes and grasses, occurying in highly-distubed and nitriphicated soils (**Fig. 4**). Wide distribution in the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Digitaria sanguinalis*, *Echinochloa crus-galli*, *Chenopodium album*, *Amaranthus hybridus* aggr.

***Chenopodietea*** Winter-annual ruderal vegetation.

10. *Allion triquetri* Geophyte-rich fringe vegetation. Occupies warmer and shadier situations than the other alliances in the class, has a higher proportion of geophytes and a lower proportion of therophytes, and is adapted to less frequent disturbances (**Fig. 4**). Wide distribution in the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Urtica membranacea*, *Smyrnium olusatrum*, *Tradescantia fluminensis*.

11. *Geranio pusilli-Anthriscion caucalidis* Winter-annual fringe vegetation. Occupies situations similar to the previous alliance, but in colder sites, and is dominated by therophytes instead of geophytes (**Fig. 4**). Mostly distributed in the more Mediterranean valleys south of the Cantabrian Mountains (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Geranium lucidum*, *Anthriscus caucalis*.

12. *Chenopodion muralis* Low-growth winter-annual ruderal vegetation. Occupies colder situations than the other alliances in the class and more often disturbed sites, and has a longer and later flowering season (**Fig. 4**). Sparse occurencce throughout the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Malva neglecta*, *Urtica urens*.

13. *Echio-Galactition tomentosae* Tall-herb winter-annual ruderal vegetation. Occupies drier, sunnier and less nutrient-rich situations than the other alliances in the class (**Fig. 4**). Mostly distributed in the west of the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Galactites tomentosus*, *Coleostephus myconis*, *Anisantha rigida*.

14. *Linario polygalifoliae-Vulpion alopecuri* Ephemeral annual vegetation of disturbed coastal dunes. Occupies drier and more nutrient-poor situations than the other alliances in the class (**Fig. 4**) and has a high number of coastal dune species. Isolated occurrence in the southern coasts of the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Vulpia alopecuros*, *Malcolmia littorea*, *Crucianella maritima*, *Artemisia campestris* aggr.

***Sisymbrietea*** Summer-annual ruderal vegetation.

15. *Sisymbrion officinalis* Summer-annual ruderal vegetation. Rich in therophytes and adapted to severe disturbances (**Fig. 4**). Wide distribution in the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Hordeum murinum*, *Sisymbrium officinale*.

***Artemisietea vulgaris*** Perennial ruderal vegetation of dry sites.

16. *Convolvulo arvensis-Agropyrion repentis* Semiruderal grasslands. Occupies wetter and shadier situations than the other alliances in the class, and has a higher proportion of geophytes and grasses (**Fig. 4**). Isolated occurencce in the south of the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Poa compressa*, *Elytrigia repens*, *Potentilla reptans*, *Valerianella locusta*.

17. *Carduo carpetani-Cirsion odontolepidis* Thistle ruderal vegetation of warm sites. Occupies warmer situations than the other thistle in the class (**Fig. 4**). Mostly distributed in the Cantabrian Mountains (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Onopordum acanthium*.

18. *Cirsion richterano-chodati* Thistle ruderal vegetation of cold sites. Occupies colder situations than the other thistle alliance in the class (**Fig. 4**). Mostly distributed in the Cantabrian Mountains (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Cirsium eriophorum*, *Carduus carpetanus*, *Carduus nutans* aggr.

19. *Dauco-Melilotion* Biennial ruderal vegetation. Occupies wetter situations than the two thistle alliances in the class, and has a higher proportion of therophytes (**Fig. 4**). Sparse occurencce throughout the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Helminthotheca echioides*, *Daucus carota*, *Melilotus albus*, *Foeniculum vulgare*.

***Epilobietea angustifolii*** Perennial ruderal vegetation of mesic to wet sites.

20. *Geo urbani-Alliarion officinalis* Low-herb short-lived semiruderal and fringe vegetation. Occupies shadier and less disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig. 4**). Wide distribution in the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Urtica dioica*.

21. *Arction lappae* Low-herb short-lived ruderal vegetation. Occupies sunnier and more frequently-disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig. 4**). Distribution concentrated in the Cantabrian Mountains (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Oxybasis rubra*, *Senecio duriaei*.

22. *Balloto-Conion maculati* Tall-herb perennial ruderal vegetation. Occupies warmer situations than the other alliances in the class (**Fig. 4**). Distribution concentrated in the Cantabrian Mountains and the east of the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Sambucus ebulus*.

23. *Aegopodion podagrariae* Tall-herb perennial vegetation of forest margins and clearings. Occupies shadier and less disturbed situations than the other alliances in the class (**Fig. 4**). Sparse occurencce throughout the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Anthriscus sylvestris*, *Heracleum sphondylium*, *Galium aparine*.

24. *Epilobion angustifolii* Tall-herb perennial vegetation of forest margins and clearings in acidic soils. Occupies colder, nutrient-poorer, more acidic and less disturbed situations than the other alliances in the class (**Fig. 4**). Distributed towards the west of the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Epilobium angustifolium*, *Digitalis purpurea*, *Luzula lactea*.

25. *Cynancho-Convolvulion sepium* Tall-herb vegetation of nutrient-rich riparian habitats. Occupies wetter situations and less disturbed situations than the other alliances in the class, and has a later flowering season (**Fig. 4**). Dispersed throughout the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Eupatorium cannabinum*, *Angelica sylvestris*, *Picris hieracioides*.

26. *Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich river banks and ditches. Occupies warmer and wetter situations situations than the other alliances in the class, flowers later, has taller plants and is highly dominated by neophytes and geophytes (**Fig. 4**). Isolated occurence in the western coast of the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Arundo donax*, *Silene latifolia*.

***Bidentetea*** Summer-annual pioneer vegetation of temporarily-flooded sites.

27. *Bidention tripartitae* Summer-annual pioneer vegetation of temporarily-flooded sites. Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes (**Fig. 4**). Dispersed throughout the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*.

28. *Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily-flooded and warm sites. Occupies warmer and wetter situations than the other alliance in the class, has a higher proportion of neophytes and has a longer flowering season (**Fig. 4**). Isolated occurrences in the south of the ecoregion (**Fig. 1**). Diagnostic species (**Supplementary 5**): *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*.
